---
title: "Demo R jump start"
format: pdf
editor: visual
author: Danyili Hong
date: '9/16/2023'

---

## Data

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggformula)
library(ggplot2)
```

First I will read my dataset:

```{r}
pp <- read_csv('https://sldr.netlify.app/data/sustainable-livelihoods.csv')
head(pp)
```

Are men or women more likely to participate in the agriculture program? Are these results different between countries?

Graphics

```{r}
gf_percents(~Gender, fill = ~Agriculture_Program, data = pp, position = 'dodge')
```

```{r}
gf_bar(~Gender | Country, fill = ~Agriculture_Program, data = pp, position = 'dodge')
```

text

I chose side-by-side bar graph because I am looking for the correlation between two categorical variables. I also used facet(Country) the third variable to see the relationship.


According to the second graph, men in Mali participated in the agriculture program the most. Women in Tanzania participated in the agriculture program the most. 

critique: 
I think I can answer questions in a more professioned way rather than report what is obvious on the graph.